---
title: "arrendajo_mirlo"
author: "Antonio J Perez-Luque"
date: "2021-06-23"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Objetivo
-  Comparar datos mensuales de abundancia de *Garrulus glandarius* y *Torquatus viscivorus* para los periodos de dispersión de semillas (septiembre, octubre y noviembre)

## Arrendado
### Prepara datos de Arrendajo por mes 

```{r, echo = FALSE, message=FALSE, warning=FALSE}
library("tidyverse")
library("here")
library("car")
library("report")
library("plotrix")
library("broom")
library("kableExtra")
library(ggstatsplot)
library(statsExpressions)
library(plotrix) #std.error 
library("flextable")
library(plotrix)
library("ggsci")
```


*Obsnev bird censuses* provided by [*OBSNEV*](https://obsnev.es/), realized in several transects distributed along Sierra Nevada. Temporal range from 2008 to 2020. The data were downloaded from new information system of OBSNEV (*i.e.* PostgreSQL [db01.obsnev.es](db01.obsnev.es))
  

- Raw data were downloaded from OBSNEV information system. The downloaded tables were: 
`contactos_paseriformes.csv`; `dicc_especies.csv`; `geo.csv`; `visitas.csv`. 

- The protocol sampling number for passerine in the database is 5. 

- Select the two locations: "Cortijo del Hornillo" (SANJUAN), "Robledal de Canar" (CANAR)

- Select only the taxonomic level = species (level > 6) 

- Filter out data contacted over 30 m 

```{r read-obsnev-tables}
contactos <- read_csv(here::here("data/data_raw/contactos_paseriformes.csv")) %>% 
  dplyr::select(-fcreacion, -fmodificacion)

dicc_sp <- read_csv(here::here("data/data_raw/dicc_especies.csv")) %>% 
  dplyr::select(idesp, nombre_cientifico, nivel)

# El protocolo de paseriformes es el 5
dicc_visita <- read_csv(here::here("data/data_raw/visitas.csv"), 
                        col_types = 
                          cols(.default ="?", 
                               idgeo = col_character(), 
                               fvisita = col_datetime(format="%Y-%m-%d %H:%M:%S"))) %>%
  filter(protocolo == 5) 
  
dicc_geo <- read_csv(here::here("data/data_raw/geo.csv"), 
                     col_types = cols(.default ="?", 
                               longitud_m = col_double()))

visita_geo <- 
  dicc_visita %>% inner_join(dicc_geo, by = "idgeo") %>% 
  dplyr::select(idvisitas, fvisita, nombre, longitud_m) 

dfraw <- contactos %>% 
  inner_join(visita_geo, by = "idvisitas") %>% 
  inner_join(dicc_sp, by = "idesp") %>% 
  mutate(year = lubridate::year(fvisita), 
         mes = lubridate::month(fvisita))

df <- dfraw %>% 
  filter(nombre %in% c("Cortijo del Hornillo", "Robledal de Cáñar")) %>% 
  mutate(habitat = case_when(
    stringr::str_detect(nombre, "Hornillo") ~ "SANJUAN",
    stringr::str_detect(nombre, "Cáñar") ~ "CANAR")) %>% 
  mutate(year = lubridate::year(fvisita), 
         mes = lubridate::month(fvisita),
         fecha = strftime(fvisita, format="%Y-%m-%d")) %>% 
  filter(nivel > 6) %>% 
  filter(desplazamiento < 31)

```


- Aggregated data to get the sum of all contacts for a specie recorded in the same visit of the same transect

- Select data from month = 9, 10, 11 

```{r, compute-abundance-obsnev}
garrulus <- df %>% 
  filter(nombre_cientifico == "Garrulus glandarius") %>% 
  group_by(fecha, year, mes, longitud_m, nombre, habitat, nivel) %>%
  summarise(total_ind = sum(numero)) %>%  
  mutate(den = round((total_ind * 10000 * 10 / (longitud_m * 60)),2)) %>% 
  ungroup() %>% 
  rename(localidad = habitat) %>% 
  dplyr::select(-nivel, -longitud_m, -nombre, -total_ind, -fecha) %>% 
  filter(mes %in% c(9,10,11)) %>% 
  filter(year > 2007)
  
dataRZ <- data.frame(year = rep(1984,3),
                    mes = c(9,10,11), 
                    localidad = rep("SANJUAN", 3),
                    den = c(2.7,1.5,1.5))

garrulus_all <- bind_rows(garrulus, dataRZ)
```


```{r}
garrulus %>% filter(localidad == "SANJUAN") %>% 
  ggplot(aes(x=year, y=den, fill=as.factor(mes))) + 
  geom_bar(stat="identity", position = position_dodge(.9))
  
garrulus_mean <- garrulus %>% 
  group_by(mes, localidad) %>% 
  summarize(den_avg = mean(den, na.rm=TRUE), 
            den_sd = sd(den, na.rm = TRUE), 
            den_se = std.error(den, na.rm=TRUE)) %>% 
  mutate(date = "2008-2018") 


dataRZ_mean <- data.frame(date = rep("1984",3),
                    mes = c(9,10,11), 
                    localidad = rep("SANJUAN", 3),
                    den_avg = c(2.7,1.5,1.5),
                    den_sd = NA, den_se=NA)


compare_garrulus <- bind_rows(garrulus_mean, dataRZ_mean)
```

Plot 

```{r}
pos <- position_dodge(0.9)

ab_garrulus <- compare_garrulus %>% 
  mutate(mes = factor(mes, levels = c("9","10","11"))) %>% 
  filter(localidad == "SANJUAN") %>% 
  ggplot(aes(x=date, y=den_avg, 
             fill= mes, color = mes)) + 
  # geom_point(position = pos) + 
  geom_bar(stat="identity", position = pos, width = .8) +
  geom_errorbar(aes(ymin=den_avg - den_se, ymax = den_avg + den_se), 
                position = position_dodge(0.9), width = .5) + 
  ylab("Abundance (Birds/10ha)") + xlab("") + 
  ylim(0,3) + 
  theme_bw() +
  scale_fill_jco() +
  scale_color_jco() +
  theme(panel.grid = element_blank(),
         plot.title = element_text(face="italic")) +
  ggtitle("Garrulus glandarius") +
  guides(col=guide_legend("month"),
         fill=guide_legend("month"))
```


## Torcuatus data 

```{r}
df_turdus <- dfraw %>% 
  mutate(year = lubridate::year(fvisita), 
         mes = lubridate::month(fvisita),
         fecha = strftime(fvisita, format="%Y-%m-%d")) %>% 
  filter(nivel > 6) %>% 
  filter(desplazamiento < 31) %>% 
  filter(nombre_cientifico == "Turdus torquatus") 

# n Campos de Otero = 22 

turdus <- df_turdus %>% 
  filter(nombre == "Campos de Otero") %>% 
  group_by(fecha, year, mes, longitud_m, nombre, nivel) %>%
  summarise(total_ind = sum(numero)) %>%  
  mutate(den = round((total_ind * 10000 * 10 / (longitud_m * 60)),2)) %>% 
  ungroup() %>% 
  dplyr::select(-nivel, -longitud_m, -nombre, -total_ind, -fecha) %>% 
  filter(mes %in% c(9,10,11)) %>% 
  filter(year > 2007) %>% 
  mutate(date = "2008-2018")


turdus_old <- data.frame(
  year = c(1984, 1984, 1984, 1985, 1985),
  mes = c(9, 10, 11, 10, 11),
  den = c(0.1, 2.6, 8.7, 3.6, 9.8),
  date = rep("1984-1985",5)
)
  
turdus_all <- bind_rows(turdus, turdus_old)  
  
turdus_avg <- turdus_all %>% 
  group_by(mes, date) %>% 
  summarize(den_avg = mean(den, na.rm=TRUE), 
            den_sd = sd(den, na.rm = TRUE), 
            den_se = std.error(den, na.rm=TRUE))
  
  
pos <- position_dodge(0.9)

ab_turdus <- turdus_avg %>% 
  mutate(mes = factor(mes, levels = c("9","10","11"))) %>% 
  ggplot(aes(x=date, y=den_avg, 
             fill= mes, color = mes)) + 
  # geom_point(position = pos) + 
  geom_bar(stat="identity", position = pos, width = .8) +
  geom_errorbar(aes(ymin=den_avg - den_se, ymax = den_avg + den_se), 
                position = position_dodge(0.9), width = .5) + 
  ylab("Abundance (Birds/10ha)") + xlab("") + 
  theme_bw() +
  scale_fill_jco() +
  scale_color_jco() +
  theme(panel.grid = element_blank(), 
        plot.title = element_text(face="italic")) +
  ggtitle("Turdus torquatus") + 
  guides(col=guide_legend("month"),
         fill=guide_legend("month"))


```


```{r}
library(patchwork)
gar_tor <- ab_turdus + ab_garrulus &
  theme(legend.position = "bottom") 
garrulus_torquatus <- gar_tor +  plot_layout(ncol=1, guides = "collect") 
```

```{r}
garrulus_torquatus
```

```{r}
ggsave(here::here("output/garrulus_torquatus.pdf"),
       width =10, height = 15, units = "cm")
gar_tor
dev.off()
```






