---
title: "explore-trends"
author: "Antonio J Perez-Luque"
date: "2021-04-29"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Explore trends 

```{r, set-global-options}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
```

```{r load-pkgs}
library("tidyverse")
library("here")
library("plotly")
```

- Read data and compute average abundance by year 

```{r, read-data}
passerine <- read_csv(here::here("data/passerine.csv")) 
passerine_abbreviations <- read_csv(here::here("data/passerine_abbrev.csv")) 

passerine.ab <- passerine %>% 
  rename(sp = sp.abb) %>% 
  group_by(especie, sp, year, habitat) %>% 
  summarise(ab_avg = round(mean(den, na.rm=TRUE),2), 
            sd = sd(den, na.rm = TRUE), 
            se = sd/sqrt(length(den)), 
            n = length(den)) 
```


```{r, define-ggplot-theme}
bird_theme <- 
  ggplot2::theme_bw() + 
  ggplot2::theme(
    panel.grid.minor = element_blank(),
    strip.background = element_blank()
  )
```

- We explore the abundance over time for each habitat, and then select those species with *n* years > 3. 

## Summit environment  

```{r, general-plot-cumbres}
cumbres.plotall <- passerine.ab %>% 
  filter(habitat == "cumbres") %>% 
  ggplot(aes(x=as.factor(year), y=ab_avg)) + 
  geom_errorbar(aes(ymin = ab_avg - se, 
                    ymax = ab_avg + se),
                width = 0.2) +
  geom_point() +
  facet_wrap(~especie) + 
  ylab("abundance (ind / 10ha)") + xlab("") + 
  bird_theme
```

```{r, plot-summits-all, echo=FALSE, message=FALSE}
ggsave(here::here("output/abundance_time/plot_all_cumbres.pdf"),
       width = 29, height = 15, units = "cm")
cumbres.plotall
dev.off()
```


- Plot only the selected taxa (contacted in more than three years)

```{r, plot-cumbres}
cumbres.plot <- passerine.ab %>% 
  filter(habitat == "cumbres") %>% 
  filter(stringr::str_detect(especie, "Card|Oena|Phoeni|collaris")) %>%
  ggplot(aes(x=as.factor(year), y=ab_avg)) + 
  geom_errorbar(aes(ymin = ab_avg - se, 
                    ymax = ab_avg + se),
                width = 0.2) +
  geom_point() +
  facet_wrap(~especie) + 
  ylab("abundance (ind / 10ha)") + xlab("") + 
  bird_theme + 
  ggtitle("Summits")
```

```{r, plot-summits-selected, echo=FALSE, message=FALSE}
ggsave(here::here("output/abundance_time/plot_selected_cumbres.pdf"),
       width = 20, height = 15, units = "cm")
cumbres.plot
dev.off()
```

## Juniper scrublands  

```{r, general-plot-enebral}
enebral.plotall <- passerine.ab %>% 
  filter(habitat == "enebral") %>% 
  ggplot(aes(x=as.factor(year), y=ab_avg)) + 
  geom_errorbar(aes(ymin = ab_avg - se, 
                    ymax = ab_avg + se),
                width = 0.2) +
  geom_point() +
  facet_wrap(~especie) + 
  ylab("abundance (ind / 10ha)") + xlab("") + 
  bird_theme +
  theme(axis.text.x = element_text(angle=90)) 
```

```{r, plot-enebral-all, echo=FALSE, message=FALSE}
ggsave(here::here("output/abundance_time/plot_all_enebral.pdf"),
       width = 29, height = 21, units = "cm")
enebral.plotall
dev.off()
```


- Plot only the selected taxa (contacted in more than three years)

```{r, plot-enebral}
enebral.plot <- passerine.ab %>% 
  filter(habitat == "enebral") %>% 
  filter(stringr::str_detect(especie, 
                             "Alauda|Anthus|cannabina|Embe|Oena|Phoeni|collaris|rubicola|conspici|Troglo")) %>%
  ggplot(aes(x=as.factor(year), y=ab_avg)) + 
  geom_errorbar(aes(ymin = ab_avg - se, 
                    ymax = ab_avg + se),
                width = 0.2) +
  geom_point() +
  facet_wrap(~especie, nrow = 5) + 
  ylab("abundance (ind / 10ha)") + xlab("") + 
  bird_theme + 
  theme(axis.text.x = element_text(angle=90)) +  
  ggtitle("Juniper")
```

```{r, plot-enebral-selected, echo=FALSE, message=FALSE}
ggsave(here::here("output/abundance_time/plot_selected_enebral.pdf"),
       width = 18, height = 21, units = "cm")
enebral.plot
dev.off()
```


## Oak populations   

```{r, general-plot-robledal}
robledal.plotall <- passerine.ab %>% 
  filter(habitat == "robledal") %>% 
  ggplot(aes(x=as.factor(year), y=ab_avg)) + 
  geom_errorbar(aes(ymin = ab_avg - se, 
                    ymax = ab_avg + se),
                width = 0.2) +
  geom_point() +
  facet_wrap(~especie) + 
  ylab("abundance (ind / 10ha)") + xlab("") + 
  bird_theme +
  theme(axis.text.x = element_text(angle=90))   
```

```{r, plot-robledal-all, echo=FALSE, message=FALSE}
ggsave(here::here("output/abundance_time/plot_all_robledal.pdf"),
       width = 29, height = 21, units = "cm")
robledal.plotall
dev.off()
```

- Plot only the selected taxa (contacted in more than three years)

```{r, plot-robledal}
robledal.plot <- passerine.ab %>% 
  filter(habitat == "robledal") %>% 
  filter(stringr::str_detect(especie, 
                             "Aeg|cannabina|Certhia|Cyanis|Erith|Fringi|Garrulus|Lopho|Lullula|Parus|Peripares|Phoeni|bonelli|Regulus|rubicola|serinus|Sitta|atricapilla|cantillans|merula|visci|Troglo")) %>%
  ggplot(aes(x=as.factor(year), y=ab_avg)) + 
  geom_errorbar(aes(ymin = ab_avg - se, 
                    ymax = ab_avg + se),
                width = 0.2) +
  geom_point() +
  facet_wrap(~especie, nrow = 5) + 
  ylab("abundance (ind / 10ha)") + xlab("") + 
  bird_theme + 
  theme(axis.text.x = element_text(angle=90)) +  
  ggtitle("Oak woodlands")
```

```{r, plot-robledal-selected, echo=FALSE, message=FALSE, warning=FALSE}
ggsave(here::here("output/abundance_time/plot_selected_robledal.pdf"),
       width = 18, height = 21, units = "cm")
robledal.plot
dev.off()
```







