---
title: "prepare-data"
author: "Antonio J Perez-Luque"
date: "2021-04-28"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---
```{r, set-global-options}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
```


## Data sources

Data coming from two sources: 

  - *Old bird censuses* provided by R. Zamora, consists in bird censuses in three locations: oak population (1700 masl); juniper-scrubland (2230 masl) and summit environments (3200 masl). Range temporal cover from 1981 to 1985. 
  
  - *Obsnev bird censuses* provided by [*OBSNEV*](https://obsnev.es/), realized in several transects distributed along Sierra Nevada. Temporal range from 2008 to 2020. The data were downloaded from new information system of OBSNEV (*i.e.* PostgreSQL [db01.obsnev.es](db01.obsnev.es))
  
All data are stored in the folder `/data_raw` 

## Prepare Old bird censuses 
### Notes for Old bird censuses

- File `RObledal año 1981 RZAves_SN_10ha.xls`: 

  - habitat: oak woodlands (*Q. pyrenaica*)
  - elevation: 1700 masl 
  - year: 1981 
  - variable: bird abundance monthly aggregated (ind / 10 ha)
  - sample size: n = 3 (may, june, july)
  - notes: The data aren't the original bird censuses but are monthly aggregated

- File `Enebral año 1985 RZAves_SN_10ha.xls`

  - habitat: juniper-scrubland
  - elevation: 2300 masl 
  - year: 1985
  - variable: bird abundance monthly aggregated (ind / 10 ha)
  - sample size: n = 3 (may, june, july)
  - notes: The data aren't the original bird censuses but are monthly aggregated

- File `Aves_SN_meses_reproduccion.xls` 

  - habitat: jseveral habitats. We selected juniper-scrubland and summit environments 
  - elevation: 2300 and 3200 masl 
  - year: 1984 (juniper); 1982 (summit environment)
  - variable: raw bird abundance. For juniper ind / 10.2 ha; and for summit ind / 20 ha. 
  - sample size: n = 9 transects (juniper) and 6 transects (juniper) during may-july
  - notes: Original bird censuses


```{r load-pkgs, message=FALSE}
library("tidyverse")
library("here")
library("readxl")
library("DT")
```


```{r oak-data}
robledal1981 <- read_excel(here::here("data/data_raw/RObledal año 1981 RZAves_SN_10ha.xls")) %>% 
  pivot_longer(cols= mayo_1981:julio_1981, names_to="fecha") %>% 
  separate(fecha, into = c("mes", "year"), sep="_", remove = FALSE) %>% 
  rename("especie" = Aves, "den" = value) %>% 
  mutate(year = as.numeric(year), 
         habitat = "robledal", 
         cota = 1700,
         mes = case_when(
           mes == "mayo" ~ as.numeric(5), 
           mes == "junio" ~ as.numeric(6),
           mes == "julio" ~ as.numeric(7)), 
         fecha = format(as.Date(paste(year, mes, "01", sep="-")), format="%Y-%m-%d")) 

head(robledal1981)
```


```{r, juniper85-data}
enebral1985 <- read_excel(here::here("data/data_raw/Enebral año 1985 RZAves_SN_10ha.xls")) %>%   pivot_longer(cols= mayo_1985:julio_1985, names_to="fecha") %>% 
  separate(fecha, into = c("mes", "year"), sep="_") %>% 
  rename("especie" = Aves, "den" = value) %>% 
  mutate(year = as.numeric(year), 
         habitat = "enebral", 
         cota = 2230, 
         mes = case_when(
           mes == "mayo" ~ as.numeric(5), 
           mes == "junio" ~ as.numeric(6),
           mes == "julio" ~ as.numeric(7)),
         fecha = format(as.Date(paste(year, mes, "01", sep="-")), format="%Y-%m-%d")) 

head(enebral1985)
```


```{r, juniper84-data}
enebral1984 <- read_excel(here::here("data/data_raw/Aves_SN_meses_reproduccion.xlsx"),
                           sheet = "2230") %>% 
  rename("especie" = Ave, "den" = `Número`) %>% 
  mutate(den = round(den*(10/10.2),2),
         habitat = "enebral", 
         cota = 2230, 
         mes = lubridate::month(Fecha), 
         year = lubridate::year(Fecha), 
         Fecha = strftime(Fecha, format="%Y-%m-%d")) %>% 
  rename(fecha = Fecha)

head(enebral1984)  
```

```{r, summit82-data}
cumbres1982 <- read_excel(here::here("data/data_raw/Aves_SN_meses_reproduccion.xlsx"),
                           sheet = "3200") %>% 
  rename("especie" = Ave, "den" = `Número`) %>% 
  mutate(den = round(den*(10/20),2),
         habitat = "cumbres", 
         cota = 3200, 
         mes = lubridate::month(Fecha), 
         year = lubridate::year(Fecha),
        Fecha = strftime(Fecha, format="%Y-%m-%d")) %>% 
  rename(fecha = Fecha)

head(cumbres1982)
```

- Bind old data

```{r bind-old-data}
old_census <- bind_rows(cumbres1982, enebral1984, enebral1985, robledal1981)
datatable(old_census)
```

## Prepare OBSNEV bird censuses 

- Raw data were downloaded from OBSNEV information system. The downloaded tables were: 
`contactos_paseriformes.csv`; `dicc_especies.csv`; `geo.csv`; `visitas.csv`. 

- The protocol sampling number for passerine in the database is 5. 

- Select the three locations: "Cortijo del Hornillo" (*oak*), "Campos de Otero" (*juniper*), and "Aguas Verdes" (*summits*). 

- Filter data only for may, june, july 

- Filter out data contacted over 30 m 

- Select only the taxonomic level = species (level > 6) 

```{r read-obsnev-tables}
contactos <- read_csv(here::here("data/data_raw/contactos_paseriformes.csv")) %>% 
  dplyr::select(-fcreacion, -fmodificacion)

dicc_sp <- read_csv(here::here("data/data_raw/dicc_especies.csv")) %>% 
  dplyr::select(idesp, nombre_cientifico, nivel)

# El protocolo de paseriformes es el 5
dicc_visita <- read_csv(here::here("data/data_raw/visitas.csv"), 
                        col_types = 
                          cols(.default ="?", 
                               idgeo = col_character(), 
                               fvisita = col_datetime(format="%Y-%m-%d %H:%M:%S"))) %>%
  filter(protocolo == 5) 
  
dicc_geo <- read_csv(here::here("data/data_raw/geo.csv"), 
                     col_types = cols(.default ="?", 
                               longitud_m = col_double()))

visita_geo <- 
  dicc_visita %>% inner_join(dicc_geo, by = "idgeo") %>% 
  dplyr::select(idvisitas, fvisita, nombre, longitud_m) 

dfraw <- contactos %>% 
  inner_join(visita_geo, by = "idvisitas") %>% 
  inner_join(dicc_sp, by = "idesp") %>% 
  mutate(year = lubridate::year(fvisita), 
         mes = lubridate::month(fvisita))

df <- dfraw %>% 
  filter(nombre %in% c("Cortijo del Hornillo", "Campos de Otero", "Aguas Verdes")) %>% 
  mutate(habitat = recode(nombre, 
                          "Campos de Otero" = "enebral",
                          "Cortijo del Hornillo" = "robledal",
                          "Aguas Verdes" = "cumbres")) %>%
  mutate(cota = case_when(
    habitat == "enebral" ~ 2230, 
    habitat == "robledal" ~ 1700,
    habitat == "cumbres" ~ 3200
  )) %>% 
  mutate(year = lubridate::year(fvisita), 
         mes = lubridate::month(fvisita),
         fecha = strftime(fvisita, format="%Y-%m-%d")) %>% 
  filter(mes %in% c(5,6,7)) %>% 
  filter(desplazamiento < 31) %>% 
  filter(nivel > 6) 

head(df)
```

- Aggregated data to get the sum of all contacts for a specie recorded in the same visit of the same transect

- Remove an error "Prunus avium" in species codification 

```{r, compute-abundance-obsnev}
dfab <- df %>% 
  filter(nombre_cientifico != "Prunus avium") %>% 
  group_by(nombre_cientifico, fecha, year, mes, longitud_m, nombre, habitat, nivel, cota) %>%
  summarise(total_ind = sum(numero)) %>%  
  mutate(den = round((total_ind * 10000 * 10 / (longitud_m * 60)),2)) %>% 
  ungroup() %>% 
  rename(especie = nombre_cientifico) %>% 
  dplyr::select(-nivel, -longitud_m, -nombre, -total_ind)
```

- Join old and new data, and export data. Be caution with this data, since it includes some species that specialists indicated to remove them. 

- Export data as `birds.csv` 

- Recode especies according to: 

  - Acanthis cannabina to Carduelis cannabina; 
  - Parus caeruleus to Cyanistes caeruleus; 
  - Parus ater to Periparus ater; 
  - Phoenicurus ochuros to Phoenicurus ochruros; 


```{r see-all-data}
birds <- bind_rows(old_census, dfab) %>% 
  mutate(especie = case_when(
    especie == "Acanthis cannabina" ~ "Carduelis cannabina",
    especie == "Parus caeruleus" ~ "Cyanistes caeruleus",
    especie == "Parus ater" ~ "Periparus ater", 
    especie == "Phoenicurus ochuros" ~ "Phoenicurus ochruros", 
    TRUE ~ especie
  ))
datatable(birds)
write_csv(birds, here::here("data/birds.csv"))
```

- Filter out species. After consultation of specialist we'll remove the following species: 

```{r, remove-sp}
remove_sp<- c("Acrocephalus dumetorum", "Alectoris rufa", "Columba palumbus", "Corvus corax", "Ficedula albicollis", "Himantopus himantopus", "Ixobrychus sturmii", "Luscinia svecica", "Monticola solitarius","Oceanodroma leucorhoa", "Prunus avium", "Puffinus yelkouan", "Pyrrhocorax pyrrhocorax")

passerine  <- birds %>% 
  filter(!especie %in% remove_sp) 

datatable(passerine)

write_csv(passerine, here::here("data/passerine.csv"))
```


  

