---
title: "Temporal beta diveristy"
author: "Antonio J Perez-Luque"
date: "2021-04-30"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

# Explore temporal beta diversity  

- We explore the temporal evolution of $\beta$-diversity (BD)

- We read about BD and found several papers: 

  - [Legrende & De Cáceres (2013)](https://doi.org/10.1111/ele.12141)
  - [Lengrende (2019)](https://doi.org/10.1002/ece3.4984)
  - [Baselga (2010)](https://doi.org/10.1111/j.1466-8238.2009.00490.x); [Baselga (2012)](https://doi.org/10.1111/j.1466-8238.2011.00756.x); [Baselga (2017)](https://doi.org/10.1111/2041-210X.12693) 
  - [Baselga et al. (2015)](https://doi.org/10.1371/journal.pone.0127913)

- There are several R packages for BD computation: betapart; adespatial. 

## Computation of Temporal Beta Diversity Index 

- For each location we explore the TBI (see Legrende 2019) in three temporal times: t0 (1982), t1 (2010), t2(2019)


```{r, set-global-options}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
```

```{r load-pkgs}
library("tidyverse")
library("here")
library("plotly")
library("adespatial")
library("DT")
library("betapart")
source(here::here("code/computeTBI.R"))
```

- Read data and compute the abundance average by year

```{r, read-data}
passerine <- read_csv(here::here("data/passerine.csv")) 
passerine_abbreviations <- read_csv(here::here("data/passerine_abbrev.csv")) 

passerine.ab <- passerine %>% 
  rename(sp = sp.abb) %>% 
  dplyr::select(-especie) %>% 
  group_by(sp, year, habitat) %>% 
  summarise(ab_avg = round(mean(den, na.rm=TRUE),2), 
            sd = sd(den, na.rm = TRUE), 
            se = sd/sqrt(length(den)), 
            n = length(den)) 
```

- Prepare data for Temporal Beta Diversity analysis:

  - generate a site.year variable = "site"."year"
  - transpose from long to wide format to get sites as rows and especies as column
  - abundance values as values
  - NA values are filled with 0

```{r, prepare-data}
df <- passerine.ab %>% 
  unite("site", habitat:year, sep="_", remove = FALSE) %>% 
  dplyr::select(-sd, -se, -n) %>% 
  pivot_wider(names_from = sp, 
              values_from = ab_avg, 
              values_fill = 0)
```

### TBI summit habitat 

- years selected: 1982, 2008, 2018

```{r, tbi-cumbres}
cumbres <- df %>% 
  filter(habitat == "cumbres") %>% 
  dplyr::select(-year) %>% 
  column_to_rownames(var="site") %>% 
  dplyr::select_if(
    negate(function(x) is.numeric(x) && sum(x) == 0)) %>% 
  arrange(year)
  

years <- df %>% filter(habitat == "cumbres") %>% arrange(year) %>% pull(year) 

years_selected <- c(1982, 2008, 2018)

tbi_summits <- computeTBI(cumbres, vector.years = years_selected, y0=1982)
```


```{r, eval=FALSE, echo=FALSE}
tbi_summits %>% 
  rename("losses" = `mean(B/den)`,
         "gains" = `mean(C/den)`,
         "D" = `mean(D)`) %>% 
  unite("pair", year_t0:year_t1, sep=" - ") %>% 
  dplyr::select(habitat:D, Change) %>% 
  pivot_longer(losses:D) %>% 
  ggplot(aes(x=pair, y=value, colour=name, group=name)) +
  geom_point() + geom_line()
```

### TBI juniper habitat 

- years selected: 1984, 2008, 2018

```{r, tbi-enebral}
enebral <- df %>% 
  filter(habitat == "enebral") %>% 
  dplyr::select(-year) %>% 
  column_to_rownames(var="site") %>% 
  dplyr::select_if(
    negate(function(x) is.numeric(x) && sum(x) == 0)) %>% 
  arrange(year)
  
years <- df %>% filter(habitat == "enebral") %>% arrange(year) %>% pull(year) 
years_selected <- c(1984, 2008, 2018)

tbi_juniper <- computeTBI(enebral, vector.years = years_selected, y0 = 1984)
```


### TBI oak habitat 

- years selected: 1981, 2008, 2018 

```{r, tbi-robledal}
robledal <- df %>% 
  filter(habitat == "robledal") %>% 
  dplyr::select(-year) %>% 
  column_to_rownames(var="site") %>% 
  dplyr::select_if(
    negate(function(x) is.numeric(x) && sum(x) == 0)) %>% 
  arrange(year)
  
years <- df %>% filter(habitat == "robledal") %>% arrange(year) %>% pull(year) 
years_selected <- c(1981, 2008, 2018)

tbi_robledal <- computeTBI(robledal, vector.years = years_selected, y0 = 1981)
```

## Explore dissimilarities 

```{r, generate-dissimilarity}
d <- bind_rows(tbi_summits, tbi_juniper, tbi_robledal) %>% 
  mutate(year_t0common = "1980's") %>% 
  rename("losses" = `mean(B/den)`,
         "gains" = `mean(C/den)`,
         "D" = `mean(D)`) %>% 
  unite("pair", year_t0common, year_t1, sep="-") %>% 
  dplyr::select(habitat:D, Change) 

write_csv(d, file=here::here("output/betadiversity/tbi.csv"))

DT::datatable(d) %>% DT::formatRound(columns = c("losses", "gains", "D"), digits=3)

```

```{r, plot-generate-dissimilarity}
cols <- c("cumbres" = "#8d96a3", "enebral" = "#66a182", "robledal" = "#edae49")
plot_BD <- d %>% 
  ggplot(aes(x=pair, fill=habitat, colour=habitat)) +
  geom_bar(aes(y=D), stat="identity", position = position_dodge2(width = .9)) + 
  ylab(expression(paste("Dissimilarity in ", beta, "-diversisty"))) +
  xlab("") + 
  theme_bw() +
  scale_color_manual(values = cols) + 
  scale_fill_manual(values = cols) + 
  geom_point(aes(x=pair, y=losses), colour="black", shape=25, 
             position = position_dodge2(width = .9)) 

ggsave(here::here("output/betadiversity/plot_betadiversity.pdf"),
       width = 12, height = 10, units = "cm")
plot_BD
dev.off()
```


### BetaPart Analysis 

```{r}
b.cumbres <- betapart.core.abund(cumbres[,-c(1:2)])
bd.cumbres <- beta.multi.abund(b.cumbres, index.family = "bray")

b.enebral <- betapart.core.abund(enebral[,-c(1:2)])
bd.enebral <- beta.multi.abund(b.enebral, index.family = "bray")

b.robledal <- betapart.core.abund(robledal[,-c(1:2)])
bd.robledal <- beta.multi.abund(b.robledal, index.family = "bray")

bd <- cbind(data.frame(robledal = unlist(bd.robledal)),
      data.frame(enebral = unlist(bd.enebral)),
      data.frame(cumbres = unlist(bd.cumbres))) %>% 
  rownames_to_column("variable") %>% 
  pivot_longer(robledal:cumbres, names_to = "habitat") 


write.csv(bd, file=here::here("output/betadiversity/betapart.csv"))

DT::datatable(bd) %>% DT::formatRound("value", digits=3)
```

